Poincaré and SimBio: a versatile and extensible Python ecosystem for modeling systems.

### 2024-04-04 article.tex
Poincaré currently limited to ODE

\usepackage{acronym}
\acrodef{CRN}[CRN]{chemical reaction network}
\acrodef{DSL}[DSL]{domain-specific language}
\acrodef{GPU}[GPU]{graphical processing unit}
\acrodef{GUI}[GUI]{graphical user interface}
\acrodef{IDE}[IDE]{integrated development environment}
\acrodef{JIT}[JIT]{just-in-time}
\acrodef{ODE}[ODE]{ordinary differential equation}
\acrodef{PyPI}[PyPI]{Python package index}
\acrodef{RHS}[RHS]{right-hand side}
\acrodef{SBML}[SBML]{Systems Biology Markup Language}
\acrodef{SDE}[SDE]{stochastic differential equation}
\acrodef{TPU}[TPU]{tensor processing unit}

@@

To simulate a system,
we created a \texttt{Simulator} instance (Figure~\ref{fig-sim})
that translates the model into the \ac{RHS} equations
and interfaces with \ac{ODE} solvers.
By default,
it creates a first-order \ac{ODE} system using \texttt{numpy} as a backend,
and uses the LSODA solver from \texttt{scipy}.
This can be easily switched to other solvers.
The Simulator wraps the output in a \texttt{pandas.DataFrame},
which can be easily plotted with the standard \texttt{plot} method.

>>>>

\usepackage{acronym}
\acrodef{CRN}[CRN]{chemical reaction network}
\acrodef{DSL}[DSL]{domain-specific language}
\acrodef{GPU}[GPU]{graphical processing unit}
\acrodef{GUI}[GUI]{graphical user interface}
\acrodef{IDE}[IDE]{integrated development environment}
\acrodef{JIT}[JIT]{just-in-time}
\acrodef{ODE}[ODE]{ordinary differential equation}
\acrodef{PyPI}[PyPI]{Python package index}
\acrodef{RHS}[RHS]{right-hand side}
\acrodef{SBML}[SBML]{Systems Biology Markup Language}
\acrodef{SDE}[SDE]{stochastic differential equation}
\acrodef{DDE}[DDE]{delay differential equation}
\acrodef{TPU}[TPU]{tensor processing unit}

@@

To simulate a system,
we created a \texttt{Simulator} instance (Figure~\ref{fig-sim})
that translates the model into the \ac{RHS} equations
and interfaces with solvers
wrapping the output in a \texttt{pandas.DataFrame},
which can be easily plotted with the standard \texttt{plot} method.
Currently,
it only supports translating into first-order \acp{ODE},
but it would be possible to add support for \acp{SDE} or \acp{DDE}.
By default,
it uses \texttt{numpy} as a backend,
and uses the LSODA solver from \texttt{scipy}.
This can be easily switched to other solvers or backends, such as \texttt{numba}.
